---
title: "Update annotations"
author: "Chirag Parsania"
date: "7/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load library
```{r , message=FALSE , warning=FALSE}
library(tidyverse)
```


## This document is about what are the changes need to be done to update the annotations in FungiExpresZ.

* Following files need to be changed
  + fungi_db_orgdb_derieved_go_data.rds
  + fungi_db_species_info.rds
  + reference_annotation_to_expr_map.rds
  
## let's have a look in to above files 

```{r}
fungi_db_orgdb_derieved_go_data <- read_rds("/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/fungi_db_orgdb_derieved_go_data.rds")
```

In the above file, change the columns `genome`,  `orgdb_cols` and `gr_cos`. Refer prepare_annotation_data_from_annotation_hub.rmd file to know more about how to update OrgDb and GRanges. The column `genome` indicates the current version of annotations and it must have unique values. 


## update genome version in existing sample info 

Before updating, Let's have look into existing sample_info.rds.

```{r}
existing_sample_info <- read_rds("/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/expression_mats_rds_files_new/sample_info.rds")
existing_sample_info %>% glimpse()
```
The column `reference_annot` need to update. It must be identical to the `genome` in the  fungi_db_orgdb_derieved_go_data.rds


## Let's update the column `reference_annot` 

```{r}
## get all old ref annot 
old_ref <- existing_sample_info$reference_annot %>% unique()

old_ref

## update fct levels 
updated_sample_info <- existing_sample_info %>% mutate(reference_annot =  fct_recode(reference_annot, 
                                              "FungiDB-42_AnidulansFGSCA4" = "FungiDB-39_AnidulansFGSCA4"  ,
                                              "FungiDB-42_CalbicansSC5314"  = "FungiDB-39_CalbicansSC5314" ,
                                              "FungiDB-42_ScerevisiaeS288c" = "FungiDB-39_ScerevisiaeS288c"))


## back up existing rds 
# write_rds(existing_sample_info ,
#           path =  "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/expression_mats_rds_files_new/__sample_info.rds")


## write new  rds 
# write_rds(updated_sample_info ,
#           path =  "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/expression_mats_rds_files_new/sample_info.rds")

```

## update reference_annotation_to_expr_map.rds file  

This file connects  species expression mat to their annotations by column `reference_annotation`. Therefore, if annotations get updated, column `reference_annotations` must get update. 

## Let's have a look into `reference_annotation_to_expr_map.rds`


```{r,eval=FALSE}
existing_annot_to_expr_map <- readr::read_rds("/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/reference_annotation_to_expr_map.rds")

existing_annot_to_expr_map 

## update column `reference_annotation` by factor levels. 
annot_to_expr_map_update <- existing_annot_to_expr_map %>% mutate(reference_annotation =  fct_recode(reference_annotation, 
                                              "FungiDB-42_AnidulansFGSCA4" = "FungiDB-39_AnidulansFGSCA4"  ,
                                              "FungiDB-42_CalbicansSC5314"  = "FungiDB-39_CalbicansSC5314" ,
                                              "FungiDB-42_ScerevisiaeS288c" = "FungiDB-39_ScerevisiaeS288c"))


## get backup of exisiting rds 
# readr::write_rds(x = existing_annot_to_expr_map , 
#                  path = "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/__reference_annotation_to_expr_map.rds" , compress = "gz")


## write new rds 
# readr::write_rds(x = annot_to_expr_map_update , 
#                  path = "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/reference_annotation_to_expr_map.rds" , compress = "gz")

```

## Add new fpkm matrix in the fungiexperesz 

```{r, eval=FALSE}
new_added_species <- c("fumigatus" ,"niger" , "auris" ,"glabrata" ,"marneffei")

existing_ref_annot  <- fungi_db_orgdb_derieved_go_data %>% dplyr::select(1,2) %>% unique() 

map_df(new_added_species , function(x){
  match_index <- grepl(x , existing_ref_annot$species) %>% which(. == TRUE)
  existing_ref_annot %>%  slice( match_index)  
})


## Crude way to add new species entry 
## 1)  get data from reference_annotation_to_expr_map.rds in to R 
## 2)  copy it to clip board 
## 3)  paste as tribble 
## 4)  add new species manually to tribble 


# existing_annot_to_expr_map <- readr::read_rds("/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/reference_annotation_to_expr_map.rds")

## bkup_existing_annot_to_expr_map 
# readr::write_rds(x = existing_annot_to_expr_map , "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/__reference_annotation_to_expr_map.rds")


updated_annot_to_expr_map <- tibble::tribble(
          ~reference_annotation,   ~expression_mat_data_file,
   "FungiDB-42_AnidulansFGSCA4",   "a_nidulans_expr_mat.rds",
   "FungiDB-42_CalbicansSC5314",   "c_albicans_expr_mat.rds",
  "FungiDB-42_ScerevisiaeS288c", "s_cerevisiae_expr_mat.rds",
  "FungiDB-42_AfumigatusAf293", "a_fumigatus_expr_mat.rds",
  "FungiDB-42_AnigerCBS513-88", "a_niger_expr_mat.rds",
  "FungiDB-42_CglabrataCBS138", "c_glabrata_expr_mat.rds",
  "FungiDB-42_TmarneffeiATCC18224", "t_marneffei_expr_mat.rds",
  "FungiDB-42_CaurisB8441", "c_auris_expr_mat.rds"
  
  )

## write updated 
# readr::write_rds(x = updated_annot_to_expr_map , "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/annotations/reference_annotation_to_expr_map.rds")


```


