---
title: "Test FungiexpresZ plots"
author: "Chirag Parsania"
date: "8/26/2019"
output: html_document
params:
  species: "Aspergillus nidulans"
  gene_expr_mat: "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/expression_mats_rds_files_new/a_nidulans_expr_mat.rds"
  sra_sample_info: "/Users/chiragparsania/Documents/Projects/1_My_PhD/9_bioseqc/fungiexpresz/app/expression_mats_rds_files_new/sample_info.rds"
  out_dir: "./test_data"
  n_random_study:  3
  n_genes: 10000
  n_samples: 10
  n_gene_groups: 4
  n_sample_groups: 3
  eval_heatmap: FALSE
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , tidy = T , message = F)
library(tidyverse)
```


## Test heatmaps for *`r params$species`*

Create a cartoon data and test them for below purposes 

+ Normal heatmap no clustering 
+ Heatmap with clustering 
+ Heatmap with gene groups 
+ Heatmap with sample groups 
+ Heatmap with column annotations (Groups)
+ Heatmap with joined data  


#### Create cartoon data  

```{r}
set.seed(1234)
cartoon_data <- matrix(rnorm(params$n_genes * params$n_samples ) , nrow = params$n_genes)
```
#### Assign column names  

```{r}
colnames(cartoon_data) <- paste("Test_col" , 1:ncol(cartoon_data) , sep = "_")
```

#### Assign row names  

Use *`r params$species`* gene names as row names. So, it can be used to test join data feature as well. 

```{r}
gene_expr_mat <-  readr::read_rds(params$gene_expr_mat)
n_random_genes <- sample(gene_expr_mat %>% pull(1) , size = params$n_genes)
rownames(cartoon_data) <- n_random_genes
```

#### Create cartoon sample groups 

```{r}
sample_groups <- factor(paste("s_grp" , 1:params$n_sample_groups,sep = "_")) %>% 
                                      sample(. , size = params$n_samples,replace = T) 
                                      
sample_grp_tbl <- tibble(groups = sample_groups , group_members = colnames(cartoon_data))

```

#### Create cartoon gene groups 

```{r}
gene_groups <- factor(paste("g_grp" , 1:params$n_gene_groups,sep = "_")) %>% 
                                      sample(. , size = params$n_genes,replace = T)
                                      
gene_grp_tbl <- tibble(groups = gene_groups , group_members = rownames(cartoon_data))
```


#### Prepare data to be joined

```{r}
sample_info <- read_rds(params$sra_sample_info) %>% filter(species == params$species)
selected_bio_project_id <- sample(sample_info$bio_project , params$n_random_study)

```

Selected bio projects are `r selected_bio_project_id`

#### Get SRA data for selected studies 

```{r}

query_sra_run_id <- sample_info %>% filter(.$bio_project %in%  selected_bio_project_id) %>% pull(run_accession)
query_sra_expr_mat  <-  gene_expr_mat %>% dplyr::select(1,query_sra_run_id)

```

#### Create joined data 
```{r}
joined_data <- cartoon_data %>% as.data.frame() %>% rownames_to_column("geneName") %>% 
  left_join(query_sra_expr_mat ) 

joined_data_mat <- joined_data %>% as.data.frame() %>% column_to_rownames("geneName") %>% as.matrix()
```

#### Save necessary data to file

```{r , eval=TRUE}
dir.create(path = params$out_dir)  

get_path_for_out_dir <- as_mapper(~paste(params$out_dir, .x , sep = "/"))

readr::write_delim(cartoon_data %>% as.data.frame() %>% rownames_to_column("geneName"), 
                   path = get_path_for_out_dir("cartoon_data.txt") , delim = "\t")
readr::write_delim(joined_data , 
                   path = get_path_for_out_dir("joined_data.txt") , delim = "\t")
readr::write_delim(sample_grp_tbl , 
                   path = get_path_for_out_dir("cartoon_sample_groups.txt") , delim = "\t")

readr::write_delim(query_sra_expr_mat , 
                   path = get_path_for_out_dir("sra_data.txt") , delim = "\t")

readr::write_delim(gene_grp_tbl ,
                   path = get_path_for_out_dir("cartoon_gene_groups.txt") , delim = "\t")

```


## Plot 1: Heatmap no cluster 

```{r,eval=FALSE}
ComplexHeatmap::Heatmap(cartoon_data ,cluster_rows = F , cluster_columns = F)
```


## Plot 2: Heatmap with row cluster 

```{r,echo=TRUE,eval=FALSE}
ComplexHeatmap::Heatmap(cartoon_data ,cluster_rows = T , cluster_columns = F)
```

## Plot 3: Heatmap with column cluster
```{r , echo=TRUE,eval=FALSE}
ComplexHeatmap::Heatmap(cartoon_data ,cluster_rows = F , cluster_columns = T)
```


## Plot 4: Heatmap with kmeans 

```{r, echo=TRUE,eval=FALSE}
set.seed(1234)
ComplexHeatmap::Heatmap(cartoon_data ,cluster_rows = F , cluster_columns = F, row_km = 3 , column_km = 3)
```


## Plot 5: Joined data heatmap

```{r,eval=FALSE}
ComplexHeatmap::Heatmap(joined_data_mat)

```

## Plot 6: Joined data heatmap with kmeans

```{r,eval=FALSE}
set.seed(1234)
ComplexHeatmap::Heatmap(joined_data_mat , column_km = 3, row_km = 3 , cluster_column_slices = F , cluster_row_slices = F)

```

## Plot 7: Heatmap with column groups top annotations 

```{r,eval=FALSE}
## elow function create default gg plot style colors
get_gg_colors <- function(x){
  #x <- letters
  colrs <- scales::hue_pal()(length(x))
  rlang::set_names(colrs , x)
}

## prepare top annot 
 hm_selected_column_groups <- joined_data_mat %>% 
          colnames() %>% 
          tibble(hm_cols = .  ) %>% 
          left_join(sample_grp_tbl , by = c(hm_cols =  "group_members")) %>%   
          pull(groups) %>%  as.character() %>% 
          replace_na("NA")
 
## color by column groups 
grp_colors <- get_gg_colors(x = hm_selected_column_groups %>% as_factor() %>% levels())

## prepare top annot         
column_ha = ComplexHeatmap::HeatmapAnnotation(Groups =  hm_selected_column_groups , col = list(Groups = grp_colors))

## heatmap 
ComplexHeatmap::Heatmap(joined_data_mat , top_annotation = column_ha , 
                        cluster_columns = F , cluster_rows = F , cluster_column_slices = F)
        
```

## Plot 8 : Heatmap split by column groups
```{r,eval=FALSE}

ComplexHeatmap::Heatmap(joined_data_mat , top_annotation = column_ha , 
                        cluster_columns = F , cluster_rows = F , cluster_column_slices = F , 
                        column_split = hm_selected_column_groups)
```

## Plot 9 : Heatmap split by gene groups 

```{r,eval=FALSE}
ComplexHeatmap::Heatmap(joined_data_mat , row_split = gene_grp_tbl$groups , 
                        cluster_columns = F, cluster_rows = F, cluster_row_slices = F)
```


## Plot 10 : Heatmap joined data split by sample groups and gene groups 

```{r,eval=FALSE}

ComplexHeatmap::Heatmap(joined_data_mat , 
                        row_split = gene_grp_tbl$groups , 
                        column_split = hm_selected_column_groups,
                        cluster_columns = T, cluster_rows = T, 
                        cluster_row_slices = F , top_annotation = column_ha)
```












